---
title: "Austin Crime Data EDA Report"
output: html_document
date: "2025-02-28"
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      warning = TRUE,
                      message = FALSE,
                      fig.align = "center",
                      R.options = list(max.print=50))
```

```{r message=FALSE, warning=FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(readr)
atxCrime <- read_csv("Crime_Reports_20250228.csv")
atxCrime
```

### 1. Dataset

Which dataset(s) are you exploring? Include the link to your dataset and quickly describe what information it contains: How many rows/columns?

```{r}
# taking a look at the atxCrime data set, and finding the number of rows and columns within the dataset.
head(atxCrime)
dim(atxCrime)
```

**We are exploring the [Austin Crime Reports](https://data.austintexas.gov/Public-Safety/Crime-Reports/fdj4-gpfu/about_data) data set which reports crime incidents that the Austin Police Department responded to and documented, from January 1st, 2003 to February 22nd, 2025. Each row in the data set represents a unique crime report, detailing various aspects of the incident, such as the highest offense, date, location type, and date/time of occurrence. The data set has 2532115 rows, indicating the number of unique crime reports, and 6 columns, indicating the 6 variables we are interested in exploring - Highest Offense Description, Family Violence, Location Type, Council District, Occurred Date/Time, and Report Time/Date**.

2\. Research Question

Write a research question involving specific variables to guide your Exploratory Data Analysis.

**We will explore this data set to answer the following research questions:**

1.  **Which district has the least and most crime of any type, and which crime type is the most common in each district, including family violence? (Sophia)**
2.  **How does time of year, crime type, and location type affect the frequency of criminal activity and the number of hours it took to report the crime? (Berke)**

### 3. Wrangling

Organize your data by using single-word names for all variables, filter your data on some conditions (if needed), and create new variables (if needed).

```{r}
newCrime<-atxCrime|>
  select(Offense = `Highest Offense Description`, 
         Date = `Occurred Date Time`,
         Location = `Location Type`,
         District = `Council District`,
         Family = `Family Violence`,
         Report = `Report Date Time`
         )|>
  mutate(Date = mdy_hm(Date))|>
  mutate(Mins = difftime(mdy_hm(Report),Date, units = "mins"))|>
  #mutate(Hours = difftime(mdy_hm(Report), mdy_hm(Date), units = "hours"))
  mutate(difference = case_when(
    Mins == 0 ~ "Zero Minutes",
    Mins < 100 & Mins>0  ~ "Under 100 Minutes",
    Mins > 100 ~ "Over 100 Minutes"
  ))|>
  filter(year(Date)>=2024)
newCrime
```

### 4. Explore 1 Numeric Variable

Include a visualization and summary statistics.

```{r}
# distribution of the time it takes for a crime to be reported after it occurs.
newCrime|>
  group_by(hour(Date))|>
  summarize(freq = n())|>
  ggplot()+
  geom_histogram(aes(x = freq), binwidth = 30)

newCrime|>
  filter(!is.na(difference))|>
  ggplot()+
  geom_bar(aes(x = difference))
#nrow(filter(newCrime, is.na(difference)))
```

### 5. Explore 1 Categorical Variable

Include a visualization and summary statistics.

```{r}
newCrime|>
  group_by(Location)|>
  summarize(freq = n())|>
  slice_max(n = 10, freq)|>
  ggplot()+
newCrime|>
  group_by(Offense)|>
  summarize(freq = n())|>
  slice_max(n = 10, freq)
```

**Questions**: Let us know if you have any questions about how to manipulate your dataset(s)!
